---
title: "JABBA-Hist"
subtitle: "Analysis and Figures"
author: 
  L.T. Kell
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Estimating $F_{MNY}$ for Atlantic Mackerel}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---


```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               cache     =TRUE, 
               cache.path="../cache/01/",
               fig.path  ="../figs/01-",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```


```{r, pkgs}
library(FLCore)
library(FLBRP)
library(FLife)
library(FLCandy)
require(ggplotFL)

require(ggplot2)

library(plyr)
library(dplyr)

theme_set(theme_bw(16))
```


# ICES Category 1 Stocks

## FLStock objects

Data are in the `icesdata` package

```{r, icesdata}
library(icesdata)

data(icesdata)
data(spp)
data(info)

ctc1903=read.csv("../inputs/NorthSea_stocks.csv")[,1:14] 
ctc1903$Discards[is.na(ctc1903$Discards)]=0

names(ctc1903)[c(3:4,6)]=c(".id","year","catch")
ctc1903=ctc1903[,c(3:4,6)]
```



### Extract time series from Cat 1 assessments

```{r, catch-ices, fig.height=10, fig.width=8}
source("C:/active/flr/icesdata/R/tseries.R")

ts=tseries(icesdata)

ggplot(ddply(ts,.(.id), transform, catch=stdz(catch)))+
  geom_line(aes(year,catch),col="blue")+
  geom_text(aes(x=1950, y=2, label=.id), 
            hjust=0, vjust=1, size=3, color="black")+
  facet_wrap(~.id,scale="free_y",ncol=4)+  
  labs(title="Catch: Cat 1 stocks",   
       x    ="Year",
       y    ="") +
  theme_minimal(16)+
  theme(legend.position="none", 
        strip.text = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank())
```

**Figure `r iFig=iFig+1; iFig`** Time series of standardised catches from ICES Cat 1 stock assessments.


```{r, ssb-ices, fig.height=10, fig.width=8}
ggplot(ddply(ts,.(.id), transform, ssb=stdz(ssb)))+
  geom_line(aes(year,ssb),col="blue")+
  facet_wrap(~.id,scale="free_y",ncol=4)+
  geom_text(aes(x=1950, y=2, label=.id), 
            hjust=0, vjust=1, size=3, color="black")+
  labs(title="SSB: Cat 1 stocks", 
       x    ="Year",
       y    ="")+
    theme_minimal(16)+
  theme(legend.position="none", 
        strip.text = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank())
```

**Figure `r iFig=iFig+1; iFig`** Standarised SSB from ICES Cat 1 stock assessments.


```{r, catch-1903}
ggplot(ddply(ctc1903,.(.id), transform, catch=stdz(catch)))+
  geom_line(aes(year,catch),col="blue")+
  geom_text(aes(x=1900, y=2, label=.id), 
            hjust=0, vjust=1, size=3, color="black") +
  geom_vline(aes(xintercept=1939),colour="red")+
  facet_wrap(~.id,scale="free_y",ncol=4)+
  theme_minimal(10)+
  theme(legend.position="none", 
        strip.text   = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())+  
  labs(title="Catch",
       x    ="Year",
       y    ="")
```

**Figure `r iFig=iFig+1; iFig`** Historical catches.

```{r, catch-summary}
ctc=transmute(ts,.id=.id,year=year,data=catch)
ctc2=rbind(cbind(What="ICES",ctc),
           cbind(What="1903",mutate(ctc1903,data=catch)[,-3]))

ggplot(ddply(ctc2,.(.id), transform, catch=stdz(data)))+
  geom_line(aes(year,catch,col=What))+
  geom_text(aes(x=1900, y=2, label=.id), 
            hjust=0, vjust=1, size=3, color="black") +
  geom_vline(aes(xintercept=1939),colour="red")+
  scale_color_manual(values=c("cyan","blue"))+
  facet_wrap(~.id,scale="free_y",ncol=4)+
  theme_minimal(10)+
  theme(legend.position="none", 
        strip.text   = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())+  
  labs(title="Catch",
       x    ="Year",
       y    ="")
```

**Figure `r iFig=iFig+1; iFig`** ICES Cat 1 assessments and historical catches.

```{r, catch-2}
maxyear=ddply(ctc,.(.id), with, data.frame(maxyear=max(year)))
ctc2   =merge(ctc2,maxyear,by=".id",all.x=TRUE)
ctc2   =subset(ctc2,is.na(maxyear)|year<=maxyear)

ctc2$What=factor(ctc2$What,levels=c("ICES","1903"))
ctc3=ctc2[do.call("order",ctc2[,c(".id","year","What")]),]
ctc3=ctc3[!duplicated(ctc3[,c(".id","year")]),]
  
ggplot(ddply(ctc3,.(.id), transform, catch=stdz(data)))+
  geom_line(aes(year,catch,col=What))+
  geom_text(aes(x=1900, y=2, label=.id), 
            hjust=0, vjust=1, size=3, color="black") +
  geom_vline(aes(xintercept=1939),colour="red")+
  scale_color_manual(values=c("blue","cyan"))+
  facet_wrap(~.id,scale="free_y",ncol=4)+
  theme_minimal(10)+
  theme(legend.position="none", 
        strip.text   = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())+  
  labs(title="Catch",
       x    ="Year",
       y    ="")
```

**Figure `r iFig=iFig+1; iFig`** Updated catch time series.


Stocks with missing catch-at-age data

```{r, catch-3}
subset(ldply(icesdata, function(x) any(is.na(catch.n(x)))),V1)[,1]
```


